Rutgers Autonomous Aircraft Team\\Technical Report\\2010 AUVSI UAS Competition

### 2010-05-25 paper/paper.tex
spel check and some additions to autopilot software

 \begin{wrapfigure}{r}{0.4\textwidth}
	\includegraphics[width=0.37\textwidth]{../images/daedalus_sandbagtest.jpg}
	\caption{Load testing the underside of the \emph{Daedalus}'s wings. Shown with 16lbs of load.}
	\label{fig:sandbagtest}
\end{wrapfigure}
The plane was carefully supported under the center section, nose, and tail while pre-weighed bags of sand were slowly placed symmetrically and simultaneously on the underside of each wing, starting from the innermost rib.  The \emph{Daedalus}'s wings supported 64 lbs under static loading. This is 260\% of the maximum take off weight, the force of a 1.5$g$ accelleration multiplied by a 1.73 safety factor.  We also rocked the plane ten degrees from side to side while fully loaded to simulate differential loads during flight. Our post-test structure inspection found no warped members or cracked joints.

@@

\subsubsection{Servo Subsystem}
We are using Pololu Micro Maestro \cite{micromaestro}, which communicates with the flight computer over USB and features 6 Servo PWM outputs. On Linux, the Micro Maestro enumerates as a \texttt{ttyACM} device. A simple serial protocol is used to update the servo PWM outputs. The Micro Maestro also features a configurable timeout and failsafe position for the servo PWM outputs. We have configured the device to bring each servo to a predefined positon (throttle closed, elevator up, rudder right, ailerons left) after a 2 second new command timeout. This ensures that if the autopilot process or entire flight computer fails to respond for 2 seconds, the aircraft will enter this mode until the safety pilot takes control.

@@

\subsubsection{Attitude and Heading Reference Sensor}

Early in the school year, we built a thermopile-based attitude (pitch and roll) sensor very similar to the sensor design used by the Paparazzi Tiny autopilot. Accompanied by a driver which allowed it to appear to the airborne software just as it would on the Tiny, this simple analog sensor helped us prototype and test the Paparazzi system. Having built a superior sensor, we keep it as a backup. It is detailed in Figure \ref{fig:hardware_conn_diagram} although we do not intend to use it for the competition.

Thanks to sensor donations from Invensense and ST Micro, we were able to build a 9 axis Attitude and Heading Reference Sensor (AHRS) from an Invensense ITG-3200 3-axis rate gyro \cite{itg3200} and an ST Micro LSM303DLH 3-axis accelerometer and magnetometer \cite{lsm303dlh}.

\subsubsection{Other Components}
\begin{itemize}
	\setlength{\itemsep}{0cm}
	\setlength{\parskip}{0cm}
	\item GPS: We selected the GS407 \cite{gs407gps}, a uBlox LEA-5H \cite{ublox_lea5h} based GPS with an active helical antenna. We picked the unit because the uBlox chipset has been used successfully with Paparazzi in the past, and provides an internal Kalman filter mode for aircraft expecting accellerations of $\pm2g$. The GPS is powered and connected to the USB hub with a Sparkfun FTDI adapter \cite{sparkfun_ftdi}. 
	\item Autopilot Radio: We selected a Maxstream XBee Pro (Series 2.5) 2.4GHz 50mW \cite{xbeepro24} radio system for the autopilot telemetry and uplink radio. It operates at the relatively slow baud rate of 19200bps. We selected this system because it is known to work well with the Paparazzi system, and it can provide a seamless point-to-point UART pipe with two transcievers. Importantly, it deals well with interference from other 2.4GHz radio connections nearby, including other XBee Pro radios, the Futaba ``FAAST'' radio control signal, and 2.4GHz 802.11g Wifi.
	\item Imaging Radio: We are using an identical pair of XBee Pro radios in a point-to-point 115200bps UART connection configuration to connect the imaging system to the ground. We have tested and found that, when operating two point-to-point connections at once, neither suffers measurably at up to 1/3 mile line of sight range. Therefore, rather than write complicated software to use both the autopilot serial communication protocol and the imaging radio serial communication protocol with the same XBee radio, we use two radio connections on the same band with little penalty. 
	\item Wifi Adapter: We use a 802.11g (wifi) adapter for convenience in the lab and at the field to program the flight computer and supervise the execution of autopilot proceses. The reliability of our wifi connection drops off significantly beyond 600ft (200m), so we do not actively use the link during flight or for any telemetry.
\end{itemize}
\subsubsection{Power}
Both the Beagleboard and the powered USB hub are provided with 5VDC via a switching regulator and a 3 cell (11.1v nominal) 1.5Ah Lithium Polymer battery. This provides, on average, 90 minutes of operation with all devices on.

\subsection{Software}
\label{sec:autopilot_software}
didn't want to be your ghost, didn't want to be anyone's ghost, didn't want to be your ghost, didn't want to be anyone's ghost

but i don't want anyone else
\begin{figure}
	\centering
	\includegraphics[width=0.9\textwidth]{../images/software_comm_diagram.pdf}
	\caption{Software communications diagram}	
	\label{fig:software_comm_diagram}
\end{figure}

@@

We selected the Sony MHS-PM5 camera as the basis of our imaging system. Marketed as a pocket 1080i video camera capable of 5 megapixel still images, we primarily selected the camera because it features a swiveling lens at one end of the body and it features Sony's LANC (Local Application Control Bus System) \cite{lanc}. Because of the swiveling lens we were able to achieve two rotational degrees of freedom while only rotating the body of the camera about one axis.  This greatly simiplified the pan-tilt design, and also allowed us to easily mount the camera primarily inside the fuselage of the aircraft, which protects it from damage during takeoff and landing, and reduces drag on the airframe.

@@

We took advantage of the MHS-PM5's swiveling lens assembly (see Figure \ref{fig:camera_frontview}, and added a servo actuator which can rotate the lens \degrees{135} back from straight forward, as shown in Figure \ref{fig:pantilt_linkdetail}. The tilt servo is a Hitech HS 77BB \cite{hitechs77bb}, and is capable of \degrees{180} rotation. Figures \ref{fig:pantilt} and \ref{fig:pantilt_linkdetail} show the camera lens tilted at \degrees{45} below the horizon.
The camera and tilt servo assembley is attached to the aircraft on a rotating mount. A thrust bearing takes lateral loads off the pan servo shaft, and stabilizes the assembly. The pan servo is a Hitech HS 785HB \cite{hitechs785}, and is capable of \degrees{1260} rotation (3.5 revolutions). Both servos are powered and controlled by the Pan-tilt control board.

>>>>

 \begin{wrapfigure}{r}{0.4\textwidth}
	\includegraphics[width=0.37\textwidth]{../images/daedalus_sandbagtest.jpg}
	\caption{Load testing the underside of the \emph{Daedalus}'s wings. Shown with 16lbs of load.}
	\label{fig:sandbagtest}
\end{wrapfigure}
The plane was carefully supported under the center section, nose, and tail while pre-weighed bags of sand were slowly placed symmetrically and simultaneously on the underside of each wing, starting from the innermost rib.  The \emph{Daedalus}'s wings supported 64 lbs under static loading. This is 260\% of the maximum take off weight, the force of a 1.5$g$ acceleration multiplied by a 1.73 safety factor.  We also rocked the plane ten degrees from side to side while fully loaded to simulate differential loads during flight. Our post-test structure inspection found no warped members or cracked joints.

@@

\subsubsection{Servo Subsystem}
We are using Pololu Micro Maestro \cite{micromaestro}, which communicates with the flight computer over USB and features 6 Servo PWM outputs. On Linux, the Micro Maestro enumerates as a \texttt{ttyACM} device. A simple serial protocol is used to update the servo PWM outputs. The Micro Maestro also features a configurable timeout and failsafe position for the servo PWM outputs. We have configured the device to bring each servo to a predefined position (throttle closed, elevator up, rudder right, ailerons left) after a 2 second new command timeout. This ensures that if the autopilot process or entire flight computer fails to respond for 2 seconds, the aircraft will enter this mode until the safety pilot takes control.

@@

\subsubsection{Attitude and Heading Reference Sensor}
\begin{wrapfigure}{l}{0.3\textwidth}
	\centering
	\rotatebox{-90}{
		\includegraphics[width=0.37\textwidth]{../images/ahrs_sensor_top.jpg}
	}
	\caption{Completed AHRS sensor}
	\label{fig:ahrs_picture}
\end{wrapfigure}
Early in the school year, we built a thermopile-based attitude (pitch and roll) sensor very similar to the sensor design used by the Paparazzi Tiny autopilot. Accompanied by a driver which allowed it to appear to the airborne software just as it would on the Tiny, this simple analog sensor helped us prototype and test the Paparazzi system. Having built a superior sensor, we keep it as a backup. It is detailed in Figure \ref{fig:hardware_conn_diagram} although we do not intend to use it for the competition.

Thanks to sensor donations from Invensense and ST Micro, we were able to build a 9 axis Attitude and Heading Reference Sensor (AHRS) from an Invensense ITG-3200 3-axis rate gyro \cite{itg3200} and an ST Micro LSM303DLH 3-axis accelerometer and magnetometer \cite{lsm303dlh}. Both the ITG-3200 and LSM303DLH have an all digital interface, and communicate with a microcontroller on a single I2C \cite{i2c} serial bus. Samples of all 9 axes (three each of gyro, accelerometer, and magnetometer) are taken and sent to the flight computer at 40Hz.

The two sensors and an AVR microcontroller, mounted on breakout boards, were hooked up using simple 0.1" pitch protoboard, as shown in Figure \ref{fig:ahrs_picture}. The sensor board is powered by the Sparkfun FTDI serial to USB adapter \cite{sparkfun_ftdi} the microcontroller uses to communicate with the flight computer.

\subsubsection{Other Components}
\begin{itemize}
	\setlength{\itemsep}{0cm}
	\setlength{\parskip}{0cm}
	\item GPS: We selected the GS407 \cite{gs407gps}, a uBlox LEA-5H \cite{ublox_lea5h} based GPS with an active helical antenna. We picked the unit because the uBlox chipset has been used successfully with Paparazzi in the past, and provides an internal Kalman filter mode for aircraft expecting accelerations of $\pm2g$. The GPS is powered and connected to the USB hub with a Sparkfun FTDI serial to USB adapter \cite{sparkfun_ftdi}. 
	\item Autopilot Radio: We selected a Maxstream XBee Pro (Series 2.5) 2.4GHz 50mW \cite{xbeepro24} radio system for the autopilot telemetry and uplink radio. It operates at the relatively slow baud rate of 19200bps. We selected this system because it is known to work well with the Paparazzi system, and it can provide a seamless point-to-point UART pipe with two transceivers. Importantly, it deals well with interference from other 2.4GHz radio connections nearby, including other XBee Pro radios, the Futaba ``FAAST'' radio control signal, and 2.4GHz 802.11g Wifi.
	\item Imaging Radio: We are using an identical pair of XBee Pro radios in a point-to-point 115200bps UART connection configuration to connect the imaging system to the ground. We have tested and found that, when operating two point-to-point connections at once, neither suffers measurably at up to 1/3 mile line of sight range. Therefore, rather than write complicated software to use both the autopilot serial communication protocol and the imaging radio serial communication protocol with the same XBee radio, we use two radio connections on the same band with little penalty. 
	\item Wifi Adapter: We use a 802.11g (wifi) adapter for convenience in the lab and at the field to program the flight computer and supervise the execution of autopilot process. The reliability of our wifi connection drops off significantly beyond 600ft (200m), so we do not actively use the link during flight or for any telemetry.
\end{itemize}
\subsubsection{Power}
Both the Beagleboard and the powered USB hub are provided with 5VDC via a switching regulator and a 3 cell (11.1v nominal) 1.5Ah Lithium Polymer battery. This provides, on average, 90 minutes of operation with all devices on.

\subsection{Autopilot Software}
\label{sec:autopilot_software}

The autopilot software is a port of the Paparazzi airborne software to Linux. The airborne software is written in C and originally assumed a bare-metal (that is, no operating system)

\begin{figure}
	\centering
	\includegraphics[width=0.9\textwidth]{../images/software_comm_diagram.pdf}
	\caption{Software communications diagram}	
	\label{fig:software_comm_diagram}
\end{figure}

@@

We selected the Sony MHS-PM5 camera as the basis of our imaging system. Marketed as a pocket 1080i video camera capable of 5 megapixel still images, we primarily selected the camera because it features a swiveling lens at one end of the body and it features Sony's LANC (Local Application Control Bus System) \cite{lanc}. Because of the swiveling lens we were able to achieve two rotational degrees of freedom while only rotating the body of the camera about one axis.  This greatly simplified the pan-tilt design, and also allowed us to easily mount the camera primarily inside the fuselage of the aircraft, which protects it from damage during takeoff and landing, and reduces drag on the airframe.

@@

We took advantage of the MHS-PM5's swiveling lens assembly (see Figure \ref{fig:camera_frontview}, and added a servo actuator which can rotate the lens \degrees{135} back from straight forward, as shown in Figure \ref{fig:pantilt_linkdetail}. The tilt servo is a Hitech HS 77BB \cite{hitechs77bb}, and is capable of \degrees{180} rotation. Figures \ref{fig:pantilt} and \ref{fig:pantilt_linkdetail} show the camera lens tilted at \degrees{45} below the horizon.
The camera and tilt servo assembly is attached to the aircraft on a rotating mount. A thrust bearing takes lateral loads off the pan servo shaft, and stabilizes the assembly. The pan servo is a Hitech HS 785HB \cite{hitechs785}, and is capable of \degrees{1260} rotation (3.5 revolutions). Both servos are powered and controlled by the Pan-tilt control board.
